### 2010-12-08 Paper_1.tex
Tom's edits
	modified:   Paper_1.tex

\documentclass[manuscript]{aastex} % one-column, double-spaced document
%\documentclass[preprint2]{aastex}  %% double-column, single-spaced document
% \documentclass[12pt,preprint]{aastex}
% \documentclass[aaspp4]{aastex}
% \documentclass[aas2pp4]{aastex}
% \documentclass[aj_pt4]{aastex}
% \documentclass[apjpt4]{aastex}
% \documentstyle[12pt,aasms4]{article}
% \documentclass[12pt]{article}
%\documentstyle[12pt]{article}
% \tighten
%
%
\newcommand{\ie}{{\it i.e.}}
\newcommand{\cf}{{\it cf.}}
\newcommand{\eg}{{\it e.g.}}
\newcommand{\etal}{{\it et al.}}
\newcommand{\rd}{{\rm d}}
\newcommand{\ag}{\mbox{ \raisebox{-.8ex}{$\stackrel{\textstyle >}{\sim}$} }}
\newcommand{\al}{\mbox{ \raisebox{-.8ex}{$\stackrel{\textstyle <}{\sim}$} }}
\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{tomsc@astro.indiana.edu}
\newcommand{\mso}{$M_{\odot}$ }
\newcommand{\rso}{$R_{\odot}$ }
\newcommand{\msoyr}{$M_{\odot}/{\rm year}$}

\usepackage{natbib}
\usepackage{color}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{grffile}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{amssymb}

\citestyle{aa}

% \slugcomment{Not to appear in Nonlearned J., 45.}

\shorttitle{Effective Disk Alphas}
\shortauthors{Michael et al.}

%
\begin{document}

\title{Convergence Studies on Mass Transport in Disks with Gravitational Instabilities.
 \\
I. The Constant Cooling Case}
\author{Scott Michael}
\affil{Astronomy Department, Indiana University, Bloomington, IN 47405}
\email{scamicha@astro.indiana.edu}

\author{Thomas Y. Steiman-Cameron}
\affil{Astronomy Department, Indiana University, Bloomington, IN 47405}
\email{tomsc@astro.indiana.edu}

\author{Richard H. Durisen }
\affil{Astronomy Department, Indiana University, Bloomington, IN 47405}
\email{durisen@astro.indiana.edu}

\and

\author{Aaron Boley}
\affil{Department of Astronomy, University of Florida, Gainesville, FL 32611}
\email{aaron.boley@gmail.com}

\begin{abstract}

We conducted a 3D hydrodynamic convergence study of a protostellar disk
subject to a constant global cooling time 
and to gravitational instabilities (GIs). 
The disk was followed during its asymptotic state, where
heating and cooling are in balance.
Four simulations were conducted, identical except for the number of azimuthal grid 
points, $l_{max} = $ 64, 128, 256 and 512, used in the calculations. 
The strengths of the azimuthal modes, gravitational torques, and the effective \citet{shakura1973} $\alpha$ 
arising from gravitational stresses were determined for each resolution.  These
were used to examine the resolution-dependence of each quantity
and to evaluate both the converged modal power distribution and the effective value of $\alpha$
as $l_{max} \rightarrow \infty$.   Power was found in all modes and 
the power summed over all modes was essentially independent of resolution.
The modal distribution of power, however, was significantly affected by resolution.  
The size of the azimuthal grid, $l_{max}$,
limits the number of modes to $l_{max}/2$.  Higher resolution simulations therefore
have more modes available to them.  Power is spread across all available modes.  As a consequence,
higher resolution simulations have a larger fraction of their total power in high-order modes.
High-order modes
manifest themselves more locally than low-order modes, therefore 
greater resolution leads to smaller gravitational stresses.
As a result, higher resolution simulations experience 
weaker gravitational torques than lower resolution simulations.   The effective  $\alpha$
also depends upon the magnitude of the stresses.  Thus
$\alpha_{eff}$ is also a function of grid resolution, decreasing as  $l_{max}$ increases.
The converged $\alpha_{eff}$ is consistent with
predictions from an analytic local theory for thin disks by \citet{gammie2001}, 
but only over many dynamic times when averaged over a substantial volume of the disk.
Results obtained in the $l_{max} = 512$ simulations differ with converged values
by less than 20\%.

\end{abstract}

%\keywords{ }

%%%%%% INTRODUCTION %%%%%%%%%%%%%%%
\section{Introduction}

Detailed three-dimensional hydrodynamic modeling of protostellar disks can provide considerable insight into the physical and thermal states of these disks.  However,  computing resources limit these studies to snapshots spanning timescales much shorter than a disk's evolutionary lifetime. The level of spatial resolution (and hence temporal resolution, due to the Courant condition) used in calculations can play a deterministic role in interpretations of models; while lower resolutions allow more extensive spatial and temporal studies, this benefit can come at the sacrifice of important physics.  Convergence studies provide a means 
of deriving important outcomes without the expense of very high resolution simulations.

In contrast to sophisticated calculations that include detailed physics, 
approaches exist that replace detailed physics with statistical parameters that encapsulate this physics.  This
allows for calculations covering much longer timeframes.  Both detailed and 
parametric
approaches can play important roles in broadening our understanding of complex nonlinear systems, provided a connection between the detailed physics and statistical parameters can be demonstrated and understood. In what follows, we perform convergence 
tests on hydrodynamical disk simulations to examine the applicability of an $\alpha$-disk 
formulation \citep{shakura1973} to  the
evolution of protostellar disks subject to gravitational instabilities.

%Here $\alpha$ is a dimensionless parameter that
%provides a measure of viscosity, without providing a detailed specification of the viscous mechanism.

Gravitational instabilities (GIs) can play an important, and at times dominant, role in driving the structural 
evolution of  protostellar disks \citep[for reviews, see][]{durisen2005, durisen2010}. {\bf (We need the Durisen 2010 (2011?)
review article reference)}
Thermal processes play the primary role in regulating the amplitude and outcome of these instabilities \citep{pickett1998,pickett2000,mejia2005,nelson1998,nelson2000a}. A disk's susceptibility to GIs can be parameterized by the Toomre $Q$-parameter \citep{toomre1981};  $Q = c_s \kappa / \pi G \Sigma$, where $c_s$ is the sound speed, $\kappa$ is the epicyclic frequency ($\sim$ the rotation frequency 
$\Omega$ in a nearly Keplerian disk), and $\Sigma$ is the disk surface mass density. For $Q \al$  1.5 to 1.7,  small density perturbations in a disk grow exponentially on a time scale comparable to the rotation period. These perturbations manifest themselves as multi-arm spirals with a predominantly trailing pattern that transports angular momentum outward by gravitational torques \citep{larson1984,boss1984,durisen1986,papaloizou1991, laughlin1994,nelson1998,pickett1998}.

Numerous studies have shown that thermal physics control GI amplitudes by a {\it balance} of heating and cooling \citep[e.g.,][]{tomley1991,tomley1994,pickett1998,pickett2000, pickett2003,gammie2001,boss2002,rice2003b,mejia2005,boley2006,boley2007, stamatellos2008,cossins2009}.
Heating by GIs tends to increase $c_s$, thus increasing $Q$. If disk heating persists, the instability will be suppressed once $Q$ becomes sufficiently large.  However, heating and cooling can reach a balance at nonlinear wave amplitude. GI-activity can sustain this balance  at a relatively constant,
but unstable, value of $Q$ \citep{paczynski1978, lin1981,goldreich1965}, and 
the disk can exist in a state of quasi-equilibrium. In this state, cooling is 
balanced by heating arising from the inward flow of matter and shocks induced by the GIs
\citep{gammie2001, lodato2004, rice2005, boley2006, cossins2009, vorobyov2010}.

Fully nonlinear hydrodynamic simulations of low-$Q$ disks show that multiple
modes can become unstable in the linear regime \citep{nelson1998, pickett1998, lodato2004,boley2006,cossins2009}.  Even if the initial growth is dominated by one mode, numerous modes usually appear in the nonlinear regime. As a consequence, if a disk achieves a quasi-steady balance between heating and cooling, referred to as the {\it asymptotic state},
then power will exist at all $m$-values resolved by a given numerical method \citep{mejia2005,boley2006}.  
In the asymptotic state, GI-active disks develop overlapping density structures of different strengths, geometries and coherence, producing strongly nonlinear turbulence that pervades the entire disk \citep[e.g.][]{pickett2003, mejia2005,boley2006}.
\citet{gammie2001} refers to this state as {\it gravitoturbulence}. Gravitoturbulence provides a possible bridge between the detailed physics of GIs and the viscous transport parameter $\alpha$.  Here convergence studies are important because 
higher order modes dissipate energy on shorter wavelengths than lower order modes.  The mix of low and high order modes available to a disk is set by the resolution of the simulation.

\citet{shakura1973} proposed turbulence as the primary source of effective viscosity in gas disks
and parameterized turbulent transport in a steady-state disk with the  parameter $\alpha$.  The source of the 
turbulence in their $\alpha$-disk formulation is unspecified. Although their original work was directed towards X-ray bright
accretion disks around black holes, this heuristic approach has proven fruitful in many astrophysical situations by allowing for analytic solutions and relatively easy numerical modeling over time-spans inaccessible to detailed modeling. In the 
$\alpha$-disk formalism, when the accretion rate $\dot{M}$ is constant in both time and radius, $\dot{M} = 3\pi\alpha c_s \Sigma H$ where $\Sigma$ is the surface density and $H$ represents
the disk scale height. The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it angular momentum transport
and heating are dominated by local processes}. Lin and Pringle (1987, 1990) were among the first to suggest
that GI-induced transport could be described within a viscous
framework.  However, the  validity of an $\alpha$-disk picture requires self-consistency of local energy dissipation. 
The locality of dissipation in GI-active disks has not been rigorously tested and remains an open question.

Several studies support the idea that GI transport in real disks is, in many important respects, an intrinsically global phenomenon and cannot be properly treated by a local $\alpha$-like prescription \citep{laughlin1996, balbus1999, lodato2005, mejia2005, boley2006, cai2008}.  This almost certainly applies to the embedded phase of both high and low-mass stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
rate \citep[e.g.]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}.  
However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in balance, i.e., disks subject to gravitoturbulence,
 it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process \citep{gammie2001, lodato2004, rice2005, cossins2009, vorobyov2010}.  In particular, 
an $\alpha$-viscosity may be consistent
with transport in GI acitve disks for disk to star mass ratio $M_d/M_* \le 0.2$ -- 0.3; GI-active disks
in systems with larger $M_d/M_*$ are dominated by lower-order modes that act in a more global sense and thus cannot be well represented by an $\alpha$-approach
\citep{lodato2004, cossins2009, vorobyov2010}.

\citet{gammie2001} derived an effective alpha for the case of a thin 2D gravitationally unstable disk where the heating due to local turbulent dissipation caused by GIs was balanced by the local cooling rate.  
Assuming a razor thin gravitationally unstable disk where the heating due to local turbulent dissipation caused by
GIs was balanced by a local cooling rate of $t_{cool}$.
Gammie found
$\alpha_{eff} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1}$,
where $\gamma$ is the two dimensional adiabatic index and $t_{cool}$ is a cooling parameter. 
The two dimensional index relates
to the three dimensional index, $\Gamma$, by 
$\gamma = (3\Gamma - 1) / (\Gamma + 1)$
for a non-self-gravitating
disk and
$\gamma = 3 - 2/\Gamma $
for a strongly self gravitating disk.   \cite{lodato2004} derived essentially the same result for their 3D study.
Motivated by
the \cite{laughlin1996}, \cite{gammie2001}, and \cite{lodato2004, lodato2005} results 
that demonstrated differences in how low and high order modes manifest themselves, Kratter et al. (2008) 
proposed $\alpha_{eff} = (\alpha_{short}^2 + \alpha_{long}^2)^{1/2}$, where $\alpha_{short}$ and $\alpha_{long}$
are functions of $Q$ and $M_d/M_*$ representing 
the different wavelength regimes expected to dominate for different values of the mass ratio and cooling rate. 
{\bf (Scott: need to add kratter2008 to the bib file.  The reference is Kratter, K.M., Matzner, C. D., Krumholz, M.R., 
2008. ApJ 681, 375)}
Fully 3D simulations \citep[e.g.][]{lodato2004, cossins2009} suggest that, at least in the case of imposed cooling, $\alpha_{\rm{eff}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.

We address these, and similar questions, using a grid-based, finite-difference 3D hydrodynamics code
with self gravity.
For computational grids of a fixed {\it physical} size,
grids with more elements provide finer meshes, and thus a more accurate solution. 
However, increased accuracy is offset by temporal sacrifices.
The number of equations solved at each timestep 
increases as the grid size is increased.  At the same time, the Courant condition 
forces shorter timesteps.  A mesh convergence study 
provides a common approach to balancing accuracy and computing resources.
In the following sections we follow the evolution of a disk subject to GIs
using four different
resolutions in the azimuthal direction.  The number of azimuthal grid points used in simulations
may be particularly important, because higher azimuthal resolution 
allows GI power to spread to higher-order modes 
that behave more locally. We test this by examining the power in various 
modes to see how they are affected by the choice of grid.
In the process, 
we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium; if so, to what level do these disks approximate the gravitoturbulence of \citet{gammie2001} and with what distribution of $\alpha_{eff}$; and to characterize how the GI cascades from low-order to high-order modes. To this end, we follow the evolution of a GI-active disk, subject to constant cooling, in its asymptotic state, where cooling and heating are essentially balanced.  Simulations are performed using four different resolutions in the azimuthal direction.

The balance of this paper is organized as follows.  Section 2 provides the details of the numerical approach and defines the models.  An analysis of the modal structure of the simulations 
is found in \S3.1, the gravitational torques arising from these modes follows in \S3.2.  Effective converged alphas are 
discussed in \S3.3.  Section 3.4 deals with the degree to which selected grid sizes lead to results consistent with
convergence.  Fragmentation is discussed in \S3.5.  A summary of results is found in \S4.

%We cannot yet follow the entire dynamical evolution of a 
%protoplanetary disk nor
%model all the radiative and particle processes of interest at once in full 3D. However, 
%by following GI simulations to their asymptotic states, when these exist, we
%can extrapolate the consequences over much longer times (e.g., Boley et al 2006).

\section{Hydrodynamical Simulations}

Simulations were conducted with the three-dimensional hydrodynamics code used by the Indiana University Hydrodynamics Group in several previous studies \citep{pickett1998, pickett2000, mejiaphd2004, mejia2005, boley2006, cai2008}.  This code uses a second-order, explicit, Eulerian scheme to solve Poisson's equation, an equation of state, and the equations of hydrodynamics in conservative form on a uniform cylindrical grid.  The code assumes mirror symmetry about the equatorial plane. It includes both self-gravity and artificial viscosity; the latter serves to 
mediate shocks. Source and flux terms \citep{norman1986} are computed separately in an explicit, second-order time integration \citep{albada1982,christodoulou1991,yangphd1992}.

We assume an equation of state $P = (\gamma -1)\epsilon$, where $P$ is the pressure, $\epsilon$ is the internal energy density, and $\gamma$ is the ratio of specific heats, given here by $\gamma = 5/3$.  The model disks are cooled by decreasing their internal energy according to the prescription $d\epsilon/dt = \epsilon / t_{cool}$, where $t_{cool}$ is the global cooling time in Outer Rotation Periods (ORP).  For the grid resolutions used here, one ORP is defined as the rotational period at radial zone 200.

%\subsection{Models}

Because this is a polytropic disk with an idealized cooling time, only the ratio of disk to stellar mass, given here by 0.153, is specified.  In what follows, quantities specified with real physical units are derived assuming a 1 $M_\odot$ central star and a 0.153  $M_\odot$ disk. In this case, the ORP is defined at $r = 33$ AU and one ORP $\approx$ 303 years.  Models have $t_{cool} = 2$ ORP. The disk has an initial surface density $\Sigma \sim r^{ -1}$, and initial inner and outer radii of 2.3 and 40 AU, respectively.

Initial conditions for the disk's structure and thermodynamic state were set using an equilibrium star plus disk model generated using a modified \citet{hachisu1986} self-consistent field (SCF) relaxation method \citep{pickett1996,pickett2003,mejiaphd2004,mejia2005,cai2006}. Here  the density and angular momentum distributions are iteratively solved, using the specified equation of state,  until convergence is achieved. This  equilibrium disk was then seeded with small 0.01\% $\delta\rho/\rho$ random perturbations to allow GIs to grow.  The resultant marginally unstable disk served as the initial disk for the simulation.

The initially unstable disk model passed through several phases of evolution \citep[see also][]{pickett2003, mejia2005}.  The initial disk is unstable to the growth of axisymmetric structures. During this axisymmetric phase, 
which lasts several ORPs, the disk cools and contracts, the contraction is small radially but dramatic in the vertical direction. Around 3-4 ORPs the instabilities begin grow to non-linear amplitudes. When this condition is met, the disk undergoes a strong burst of GIs that predominantly manifest themselves in one or two discrete global spiral modes.  The disk expands violently during the burst phase, producing a significant rearrangement of the disk's mass distribution on a timescale of a few ORPs. This is then followed by a period of several ORPs where heating temporarily washes out some of the nonaxisymmetry in the disk.  Finally the disk settles into a quasi-steady, long-lived asymptotic state of sustained GI-activity over a large part of the disk, with an overall balance of heating and cooling.

Simulations were run with four different angular resolutions, with grid sizes in $(r,\phi,z)$ of  $(j_{max}, k_{max}, l_{max}) =$ (512,64,64), (512,64,128), (512,64,256), and (512,64,512). Because of the large computational overhead associated with starting each simulation from $t = 0$, the $l_{max} =$ 128 simulation was run through the axisymmetric, burst and transition phases. Simulations using $l_{max} =$ 64, 256, and 512 all begin by interpolating the $l_{max} =$ 128 simulation to higher or lower azimuthal resolution at 9.6 ORPs, near the end of the transition
phase. Although this strategy saves a large amount of computational time, it limits the scope of the analysis to the asymptotic phase and it introduces a perturbation to the disk which takes several ORPs to subside. By 12 ORPs, the perturbation has dissipated and all four disks have transitioned to the asymptotic state.  All simulations were 
run through $\sim 18$ ORPs.

Figure 
%\ref{qcompareres}
\ref{fig:Final_Q} 
shows $Q$, as a function of radius, at $t \sim$ 18 ORP.  Table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of for the asymptotic disk, time-averaged of 12 -- 19 ORP and spatially averaged over 10 -- 40 AU, as a function of radius for each resolution.  From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject to the growth of nonaxisymmetric structures over this full range.  Interior to $\sim$ 27 AU, the value of $Q$ shows no dependence on
the azimuthal resolution.  From 27 -- 50 AU, lower resolution simulations display lower values of $Q$.

%%%%%%%%%%%% RESULTS & DISCUSSION %%%%%%%%%%%%%%%%

\section{Results and Discussion}

In the asymptotic state, GI-active disks develop complex density structures arising from the superposition of
all modes present in the disk.  Since the number of modes accessible to the disk is a function of the azimuthal resolution, we expect that higher resolution simulations will manifest more power in higher-order modes 
that behave more locally. We test this by examining the power in various modes to see how they are affected by the choice of grid.

\subsection{Modes}

Figure \ref{fig:DensityPlots}, shows midplane and radial densities  at $t \sim 18$ ORPs for the $l_{max} =$ 64, 128, 256, and 512 simulations. Two features are readily apparent; disks run at higher resolution are thicker than lower resolution disks {\bf (anyone want to comment on this?)} and the character of density perturbations differ with resolution.
The lower resolution (smaller $l_{max}$) simulations display longer wavelength spiral structures, with larger amplitudes and more coherence than those seen in higher resolution simulations. Higher angular resolution allows GI power to spread to higher-order azimuthal modes, as demonstrated by a Fourier decomposition of the density. The strengths of these modes are given by their global Fourier amplitudes,

\begin{equation}
A_m = \frac{(a_m^2 + b_m^2)^{1/2}}{\pi\int\rho_o \varpi {\rm d}\varpi {\rm d}z},
\end{equation}
where
\begin{subequations}
\begin{align}
a_m &= \int \rho \cos(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi,\\
b_m &=\int \rho \sin(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi.
\end{align}
\end{subequations}

Here  $\rho_o$ is the axisymmetric component of the density. In what follows, $A_m$ is averaged over time to suppress fluctuations on the dynamic time scale due to the chaotic nature of GIs in the asymptotic  phase.
In particular, $\langle A_{tot} \rangle$ will represent the power over the radial range 10 -- 40 AU, integrated from $ m = 2$ to $l_{max}/2$ and averaged over the time range 12 -- 19 ORPs. The Fourier amplitudes cannot be accurately measured at resolutions smaller than $l_{max}/2$ \citep{shannon1984}, therefore $A_m$ is limited to modes $m \le l_{max}/2$.  
Because the central star is artificially fixed to the grid center, the dynamics of the $m=1$ mode may not
be accurately treated. For this reason, it is excluded from $\langle A_{tot} \rangle$.

{\bf (Note to all: Throughout the paper I use integrals in places where we 
formally have used summations.  I note that this is a common practice, though
not strictly correct.  I find the use of integrals more elegant, but am willing to
revert to summations if people push for it.)}

Previous studies have shown that disks in the asymptotic state have power at all resolvable $m$-values  \citep{lodato2004,mejia2005,boley2006,cossins2009}.  Table \ref{tbl:ams} lists $\langle A_{tot} \rangle$ along with $\langle A_{2-7} \rangle  /  \langle A_{tot} \rangle$,  the fraction of the total power falling into low-order ($m = 2$ -- 7) modes, for each resolution. While the total power is approximately independent of resolution, $\langle A_{tot} \rangle = 1.90 \pm 0.07$, 
the distribution of this power shows marked differences. In particular, the fractional power in lower order modes decreases monotonically with increasing resolution.  Over three quarters of $\langle A_{tot} \rangle$ resides
in modes $m =$ 2 -- 7 for the $l_{max} = 64$ simulation, while approximately half resides in these modes when $l_{max} =$ 512.

The shift of power from lower to higher order modes as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$,  the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution. Thus power is naturally spread from lower order to higher order modes as the available $m$-values increases. Indeed, even if the disk is dominated by one or two low-order modes at the time of outburst, this power cascades to higher-order modes \citep[for example][]{laughlin1997, laughlin1998,laughlin1996}.

Given the resolution dependence of how power is distributed,  we must examine what resolution is sufficient to properly capture the evolution of these disks. Table \ref{tbl:ams} shows that while the fractional power in lower order modes decreases with each doubling of $l_{max}$, the amount of change in going from $l_{max} = 256$ to 512 
is significantly less than the change going from $l_{max} = 64$ to 128, with a strong suggestion that this diminution will not proceed much further with higher azimuthal resolution. Indeed, a  Richardson extrapolation \citep{press1992} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.42 for increasingly higher resolutions. This is displayed graphically in Figure \ref{fig:Power_Extrap}, where the fractional power in lower order modes is plotted against the  normalized grid 
spacing for each value of $l_{max}$. The grid spacing  is normalized such that spacing of the $l_{max}$ = 512 grid
equals unity.  Convergence occurs as the normalized spacing goes to zero. The Richardson extrapolation value  of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.42$ is plotted at a normalized grid spacing  equal to zero. The majority of the modal power in a converged disk resides in higher-order modes. 
%These modes dissipate on shorter wavelengths than low-order modes and hence act more locally.

\subsection{Gravitational Torques}

Low-order spiral modes in $\Delta\rho / \rho$  have longer wavelengths that, if coherent in radius, have longer lever 
arms to produce torques, while high-order modes have relatively short wavelengths which, particularly if they 
lack coherence, produce more localized effects. High-order modes tend to produce pockets of local GI activity, 
which can cancel each other over the whole disk.  The finding that the fractional power in high-order modes increases
with resolution, while the total power stays the same, leads to
 the expectation that  gravitational torques in the disk will decrease as $l_{max}$ is increased. This can be tested by examining the torque contributions from  various Fourier modes.

The torque {\bf C} on a cylindrical surface of the disk at radius $\varpi$ can be obtained by integrating the stress tensor $T$ over the surface of the cylinder \citep{lyndenbell1972}, i.e.,
\begin{equation}
{\rm {\bf C}} = \int {\bf r} \times T \cdot {\rm d}S.
\end{equation}
If the stress tensor includes only gravitational stresses, the surface integral can be replaced with the volume integral
\begin{equation}
{\rm \bf C} = \int \rho {\bf r} \times \nabla \Phi {\rm d}V,
\end{equation}
where $\Phi$ is the gravitational potential.  Here we are interested only in the $z$-component of torque,  
\begin{equation}
\label{torque1}
{\rm \bf C}_Z = \int \rho \frac{\partial\Phi} {\partial \phi} {\rm d}V,
\end{equation}
as only this component drives mass and angular momentum transfer. The torque contribution from each mode $m$ can be calculated by replacing $\rho$ in Equation (5) with the density distribution reconstructed from a single Fourier component
\begin{equation}
\rho_m = a_{\phi m} \cos(m\phi) + b_{\phi m} \sin(m\phi),
\end{equation}
where $a_{\phi m} =\slantfrac{1}{\pi} \int \rho \cos(m\phi){\rm d}\phi$ and $b_{\phi m} =\slantfrac{1}{\pi} \int \rho \sin(m\phi){\rm d}\phi$, and only the gravitational potential produced by the mass distribution given by $\rho_m$ is included in $\Phi$.

Figure 5  displays 
 time-averaged torques summed over a number of low-order modes, $\sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, $n = $ 1, 2, 3, 4, and 5, and $\langle {\rm \bf C}_{Ztot} \rangle$,
the total time-averaged gravitational torque arising from modes 1 thru $l_{max}/2$, for four different time-frames,
as a function of radius and azimuthal resolution. 
The scale of the $y$-axis in each subplot is the same, with the exception of the $l_{max} = 64$ simulation, 
where the scale is doubled.  
Because the star's position is artificially fixed at the center of the computational grid, 
torques arising from the $m=1$ mode,
typically 5-10\% of the total torque when averaged over the radial range of the figure,
should not be considered realistic.  
Michael \& Durisen (2010) found that freeing the star makes only 10-20\%
differences,  thus the difference between torques arising from the $m=1$ mode with the star fixed are roughly half that found when the star is free to move {\bf (Scott or Dick: please rework this sentence)}.

We note two results of Figure 5: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes for all azimuthal resolutions. Both reflect the fact that while a higher fraction of the total power in larger $l_{max}$ simulations is in higher order modes, these modes make only minimal contributions to the gravitational torque. As resolution increases, non-axisymmetric power is shifted from low-order global modes, which produce relatively large gravitational torques, to high-order local modes, which tend to cancel each other when integrated over the whole disk.

{\bf (Note to all: Figure 5 shows torques
time-averaged over for four different time-frames.  I created multiple frames because I wanted to see if torques were 
diminishing with time, as might be expected if the disk has not achieved its asymptotic state.  
Dick has told me he only believes the 12-18 ORP result - Dick, I hope 
I'm not misquoting you. Should we go with multiple panels representing different time-frames or 
go solely with the 12 - 18 orp panel?
}

\subsection{Effective Alphas}

An $\alpha$-framework treats disks as essentially two-dimensional objects subject to local energy dissipation.  Therefore, in order to compare our results with predicted and published results, we must extract two-dimensional information from our three-dimensional disks.  Following \cite{gammie2001} and \citet{lodato2004}, 
\begin{equation}
\label{alpha1}
\alpha_{\rm{eff}}(\varpi) = \left| \frac{{\rm d} \ln \Omega}{{\rm d} \ln \varpi} \right|^{-1} 
\frac {\langle T_{\varpi\phi}^{grav}\rangle + \langle T_{\varpi\phi}^{Reyn}\rangle}
{\Sigma c_s^2 },
\end{equation}
where $\Omega$ is the azimuthally averaged rotation speed, $T_{\varpi\phi}^{grav}$ and $T_{\varpi\phi}^{Reyn}$ represent the azimuthally averaged $\varpi,\phi$-component of the specific (per unit mass) gravitational 
(or ``Newton'') and hydrodynamic (or ``Reynolds'') stress tensors, respectively, $\Sigma$ the surface density,
and $c_s$ the azimuthally averaged midplane sound speed.  Brackets indicate quantities integrated over the $z$-direction. 
{\bf (All: In the previous section, brackets were used to denote time averages (torques),  while here brackets are used to denote spatial
averages.  I welcome suggestions for recasting the notation in either case)}
The denominator $\Sigma c_c^2$ was evaluated as $\int_z \rho c_c^2 {\rm d}z$. {\bf SM: I need to rewrite this entire paragraph}

%Gammie found
%\begin{equation}
%\alpha_{eff} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1}
%\end{equation}
%where $\gamma$ is the two dimensional adiabatic index and $t_{cool}$ is a cooling parameter. 
%The two dimensional index relates
%to the three dimensional index, $\Gamma$, by 
%$\gamma = (3\Gamma - 1) / (\Gamma + 1)$
%for a non-self-gravitating
%disk and
%$\gamma = 3 - 2/\Gamma $
%for a strongly self gravitating disk.

Numerical uncertainties render Reynolds stresses difficult to determine accurately. 
This difficulty is mitigated by the fact that 
previous studies have found Reynolds stresses to be small relative to gravitational stresses -- see, 
for example, Figure 5 of \cite{lodato2004} and Figure 10 of \cite{boley2006} -- suggesting that this term
can be safely neglected in the evaluation of $\alpha_{eff}$.   For this reason, 
here we omit
$\langle T_{\varpi\phi}^{Reyn}\rangle$ from the calculation of $\alpha_{eff}$. 
We note that \cite{gammie2001}  found
Reynolds and gravitational stresses contributed equally to $\alpha_{eff}$ in his study.  However,
his was a local, shearing-sheet, 2D study.  As such, it was unable to capture global effects, such
as those seen in the cited 3D studies.

The Newton stresses of equation \ref{alpha1} can be evaluated using the gravitational torque of equation \ref{torque1}, i.e.,
\begin{equation}
\label{newtonstress}
\begin{split}
 T_{\varpi\phi}^{grav}\left( \varpi \right) & = \frac{1}{2\pi\varpi^2}C_Z\\
                                & = - \frac{1}{2\pi\varpi^2} \int \rho\frac{\partial\Phi}{\partial\phi} {\rm d}V.
\end{split}
\end{equation}
Figure 6 shows $\alpha_{eff}$, time-averaged from 12 to 18 ORPs, calculated using equations \ref{alpha1} and \ref{newtonstress},
for the four grid resolutions.  For comparison, 
$\alpha_{eff}$ predicted by \citet{gammie2001} for disks where the local balance of heating and cooling dominates the energetics.  The lower curve of Gammie applies for the strongly self-gravitating limit while the upper curve 
is for the non-self-gravitating limit.   The $l_{max} = 512$ curve is consistent with these two limits.  However, 
the  $l_{max} = 512$ curve does not represent a converged model.
To determine whether the converged value of $\alpha_{eff}$ is consistent with predictions, we examine 
the spatially (10-40 AU) and temporally (12-19 ORP) averaged  $\alpha_{avg}$,
provided in Table 1, for each value of  $l_{max}$.    A Richardson extrapolation applied to these values  
converges to $\alpha_{avg} = 0.017$ (over the same radial  and time range)
as the grid spacing goes to zero.   This falls
between the $\alpha_{avg}$ values for the Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating),
with better agreement towards the strongly self-gravitating limit.

At first glance, these results suggest strong agreement with Gammie and thus strong support for dissipative
locality.  However, a cautionary note is in order.  At any given radius, 
stresses, and hence $\alpha_{eff}$, are highly variable with time.
Figure 7 shows the time-variability of $\alpha_{eff}$, 
normalized to its average value between at $t = 14$ and 18 ORP, at seven radial locations.   
Also shown is the orbital period at each radius, provided by the length of the line at the end of each subplot. 
Fluctuations in $\alpha_{eff}$
on the order of 100\% or more are seen at all radii
on  time scales range from fractions to integral multiples of the dynamic (orbital) time.  The spatial averaging
of Figure 6 washes out this local variabilty.
Thus,
even in this highly GI-active region, the predictions of Gammie are only valid over many dynamic times 
when averaged over a substantial volume of the disk.

\subsection{Convergence and $l_{max}$}

The results above permit an assessment of appropriate azimuthal grids for the
problem under consideration and similar problems.  The $l_{max} = 512$ simulation yields a value 
for $\alpha_{avg}$ of 0.020.  This  overestimates the converged value of $\alpha_{avg}
= 0.017$ by $\sim 18$\%.  The $l_{max} = 512$ simulation also overestimates the fractional
power in the high-order ($m = 2$ -- 7) modes by essentially the same amount.
Both the gravitational torques and effective alphas are functions of the
gravitational stresses within the disk.  If the Reynolds stresses are, in fact, negligible relative to the gravitational stresses,
then the errors in the calculated torques and effective alphas that stem from limited azimuthal resolution 
should be the same.   In this case, the  $l_{max} = 512$ simulation should also be expected to yield
gravitational torques that are $\sim 18$\% larger than the converged torques.  For comparison, the
errors in the $l_{max} = 256$ simulations would exceed 50\%.   For many (most?) problems of interest,
results obtained using $l_{max} = 512$ should provide reasonable agreement with results obtained 
with higher resolution.

\subsection{Fragmentation and Effective Alphas}

\noindent
{\bf (Aaron, this section is for you to write or toss, your decision.  The current text should be thought of as a filler only)}

In an $\alpha$ disk formalism the mechanism giving rise to the transport of angular momentum also acts to dissipate energy. Lodato \& Rice (2005) postulated that a disk in a state of gravitoturbulence would have locally balanced heating and cooling rates. If the heating is due to local dissipation, then $\alpha$ and $t_{cool}$ are related by
\begin{equation}
\alpha = \left| \frac{{\rm d}\ln \Omega}{{\rm d} \ln \varpi} \right|^{-2}
\frac{1}{\gamma(\gamma - 1) t_{cool}\Omega}.
\end{equation}
\citet{rice2005} concluded that for a given $t_{cool}$  there exists some maximum stress, quantified by $\alpha_{max}$, which if exceeded would cause a disk to fragment. Their predicted value $\alpha_{max} \approx 0.06$ is shown as a horizontal dotted line in Figure 6.  Note that both the $l_{max} = $ 64 and 128 simulations exceed this limiting value over some radial range, yet neither displayed fragmentation. For cases where $t_{cool}$  varies with time, \citet{clarke2007} revised this limit to $\alpha_{max} = 0.12$.

\section{Summary}

\noindent
{\bf (All: My style of writing papers is to write the summary after everything is completed.  Therefore, I have left this
section blank.)}

\acknowledgements
Please add any acknowledgements here.

\bibliographystyle{apj}
\bibliography{general}

\newpage

% \begin{deluxetable}{cccc} 
% \tablecolumns{4} 
% \tablewidth{0pc} 
% \tablecaption{Effective Alphas} 
% \tablehead{ 
% \colhead{}    &  \multicolumn{3}{c}{Non-shell Stars} &   \colhead{}   & 
%\multicolumn{3}{c}{Shell Stars} \\ 
%\cline{2-4} \cline{6-8} \\ 
%\cline{2-8} \\ 
% \colhead{Azimuthal Zones} & \colhead{$\alpha$}   & \colhead{$A_{tot}$}    
% & \colhead{$A_{2-5} / A_{tot}$} \\ 
%\colhead{} & \colhead{} &\colhead{} &  \colhead{angle} & \colhead{$10^{-28}$ ergs /}  \\
%\colhead{} & \colhead{} &\colhead{} &  \colhead{angle} 
%&  \multicolumn{2}{c}{$(10^{-28}$ ergs / cm$^3$ s sr)}  \\
% } 
% \startdata
% 64	 &    0.0587	 &        2.02 	 & 0.752	      \\
% 128	   &    0.0402	  &       1.99  	 & 0.608      \\
% 256	  &     0.0237	 &        1.87	 & 0.535     \\
% 512	  &     0.0241	 &        2.28	 & 0.467     \\
% \enddata 
% \tablecomments{
% Table 1. The $\alpha$ of figure 1, averaged over the radial region from 10 -- 40 AU, tabulated for different 
% azimuthal resolutions, along with amplitudes of Fourier components for non-axisymmetric density structures. 
% The radially averaged $\alpha$ predicted by Gammie (2001) are tablulated for comparison.
% }
% \end{deluxetable}

\begin{deluxetable}{cccccc}
\tablecolumns{5} 
\tablewidth{0pc} 
\tablecaption{$Q$, Fourier Amplitudes, and $\alpha$} 
\tablehead{ 
   \colhead{$l_{max}$} 
& \colhead{$Q_{avg}$}   
& \colhead{$\langle A_{tot} \rangle$}    
& \colhead{$\langle A_{2-7} \rangle / \langle A_{tot} \rangle$} 
& \colhead{$\alpha_{avg}$} \\ 
} 
\startdata
64   & 1.26 & 1.96 &  0.76 & 0.069 \\
128 & 1.24 & 1.91 &  0.62 & 0.044 \\
256 & 1.39 & 1.83 &  0.54 & 0.026 \\
512 & 1.48 & 1.95 &  0.49 & 0.020 \\
\enddata 
\tablecomments{
All quantities are time-averaged over 12 -- 19 ORP.
$Q_{avg}$ and $\alpha_{avg}$ are also spatially
averaged over 10 - 40 AU.
}
\label{tbl:ams}
\end{deluxetable}
\newpage

\begin{figure}
%\plotone{figures/qcompareres.eps}
\plotone{figures/qcompareres.pdf}
\caption{Toomre Q distribution versus radius for each of the four resolutions at t $\sim$ 18 ORPs}
\label{fig:Final_Q}
\end{figure}

\begin{figure}
%\plotone{figures/Figure_1.eps}
\plotone{figures/Figure_1.pdf}
\caption
{
Midplane disk densities at $t \sim 18$ ORPs for simulations with 64, 128, 256 and 512 azimuthal grid elements.
The simulations differ only in the number of  azimuthal zones used in the calculations.  Densities out of the plane, along an azimuthal cut through the disk, are displayed below the face-on views.  The color scale is  logarithmic, and axis units are AU. 
}
\label{fig:DensityPlots}
\end{figure}
\newpage

\begin{figure}
%\plotone{figures/Figure_3.eps}
\plotone{figures/Figure_3.pdf}
\caption
{
The logarithm of the amplitudes for the global Fourier components of the non-axisymmetric density structure plotted for different $m$-values. Notice that the components with the highest global amplitudes are those with low $m$-values.  
}
\label{fig:Am_vs_log_m}
\end{figure}

\begin{figure}
%\plotone{figures/Power_Extrapolation.eps}
\plotone{figures/Power_Extrapolation.pdf}
\caption
{
Convergence of fractional power.  The $x$-axis represents
the azimuthal ``spacing'' between grid points,  normalized such that spacing of the $l_{max}$ = 512 grid
equals unity.
The $y$-axis displays the fraction of the total power in the low-order ($m = $ 2--7) Fourier modes. 
Also plotted is the Richardson Extrapolation for the asymptotic value  
$\langle A_{2-7} \rangle / \langle A_{tot} \rangle = 0.42$ as the grid spacing goes to zero.  
}
\label{fig:Power_Extrap}
\end{figure}

\begin{figure}
%\plotone{figures/Torque_Array.eps}
\plotone{figures/Torque_Array.pdf}
\caption
{
Time-averaged torques summed over a number of low-order modes, 
$\sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, where $n = $1,  2, 3, 4, and 5, along
with the total time-averaged gravitational torque,
$\langle {\rm \bf C}_{Ztot} \rangle = \sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, where $n =  l_{max}/2$,
as a function of radius and azimuthal resolution. 
The scale of the $y$-axis in each subplot is the same, with the exception of the $l_{max} = 64$ simulation, 
where the scale is doubled.  
Note that the total gravitational torque is smaller for simulations run at higher azimuthal resolutions, and the total torque is dominated by low-order modes for all azimuthal resolutions. 
}
\label{fig:Torque_array}
\end{figure}

%\begin{figure}
%\plotone{figures/Mode_CascadeRatios.pdf}
%\caption
%{
%Mode Cascade Ratios .  Note that by the end of the time frame shown, there is rough parity between 
%torques arising from tower order ($m = 2$ -- 5) mades and torques from higher order modes.}
%\end{figure}

\begin{figure}
%\plotone{figures/Figure_2.eps}
\plotone{figures/Figure_2.pdf}
\caption
{Effective $\alpha$'s, time-averaged between 12 and 18 ORPs, for simulations performed with four different azimuthal grid resolutions. Dashed lines show values of $\alpha$ predicted by  \citet{gammie2001}  for the
strongly self-gravitating limit (lower line) and the non-self-gravitating limit
(upper line).
The dotted line at $\log(\alpha) = -1.22$ represents the maximum stress a disk subject to constant cooling
can maintain without fragmenting, as  proposed by  \cite{rice2005}.
}
\label{fig:alpha_v_radius}
\end{figure}
\newpage

\begin{figure}
%\plotone{figures/Alpha_variations_stack512.eps}
\plotone{figures/Alpha_variations_stack512.pdf}
\caption
{
Time variability of $\alpha$ as a function of radius.  Here $\alpha$ is normalized to its averaged value between at $t = 14$ and 18 ORP.  The orbital period at each radius is provided by the length of the bracketed line shown above its radius label. 
}
\label{fig:alpha_stack}
\end{figure}

\end{document}